The zero--one laws of Kolmogorov and Hewitt--Savage in categorical probability

### 2020-05-04 categorical-zero-one.tex
ring example

\begin{example}
$\CRing_{+}^{\op}$ is \emph{not} causal.
To see this, consider the ring $\mathbb{Z}[\epsilon]/(\epsilon^{2})$.
Elements have the form $a + b\epsilon$, with $a,b\in\bZ$, and multiplication is
defined by the convention $\epsilon^{2} = 0$.
Consider the additive maps given by
\begin{align*}
	f(a + b\epsilon) = {}& b\\
	g(a + b \epsilon) = {}& b\\
	h_{1}(a+ b \epsilon) = {}& a + b \epsilon\\
	h_{2}(a + b \epsilon) = {}& 0
\end{align*}
Since $gf = 0$, it's clear that the hypothesis holds, so if $\CRing_{+}^{\op}$
were causal, we would have the equation
$g(f(h_{1}(x)y)z) = g(f(h_{2}(x)y)z)$ for all
$x,y,z \in \bZ[\epsilon]/(\epsilon^{2})$.
However, this clearly fails if $z = \epsilon$, $y=1$ and $x = \epsilon$.
\todo[inline]{It would be nice to have such an example. But this one doesn't work, since these additive maps are not unit-preserving.}
\end{example}

>>>>

\begin{example}
$\CRing_{+}^{\op}$ is \emph{not} causal.
To see this, consider the ring $\mathbb{Z}[t]$, where the following counterexample similar to~\cite[Example~11.33]{markov_cats} can be formulated.
Consider the additive maps given as the additive extensions of
\begin{align*}
	f(t^{n}) = {}& \begin{cases}t^{n-1} & n \geq 1 \\ 1 & n = 0 \end{cases}\\[3pt]
	g(t^{n}) = {}& \begin{cases}t & n \geq 1 \\ 1 & n= 0\end{cases}\\[3pt]
	h_{1}(t^{n}) = {}& t^{n}\\[3pt]
	h_{2}(t^{n}) = {}& 1
\end{align*}
On the basis monomials $t^n$, we have $(fg)(t^n) = 1$, and therefore
\[
	(fg)(h_1(t^n)t^m) = 1 = (fg)(h_2(t^n)t^m)
\]
for all $n,m \in \mathbb{N}$, so that the hypothesis holds.
If $\CRing_{+}^{\op}$ were causal, then we would also have
\[
	f(g(h_{1}(t^n) t^m) t^\ell) = f(g(h_{2}(t^n) t^m) t^\ell)
\]
for all $n,m,\ell \in \mathbb{N}$. But this clearly fails for $n = \ell = 1$ and $m = 0$.
\end{example}
